Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-03-06 src/tex/ms.tex
order Appendices by order of appearence in main text
Section -> Appendix

We \rev{first} follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against \rev{the UV Threshold Hypothesis}.
Under \rev{this hypothesis} ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$, if \gls{NUV} exceeds the threshold for a sufficiently long period of time \rev{$\Delta T_\mathrm{min}$.}
Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.

@@

Our simulations find no clear criterion for the credibility of a biosignature detection based on spectral type of the host star, as both FGK and M~dwarf samples show similar maximum past \gls{NUV} flux distributions.
\rev{While there is a significant preference for M~dwarfs in the case of a long timescale for abiogenesis (see Section \ref{sec:long-origins}
), the fraction of inhabited planets in the M~dwarf sample drops at similar threshold fluxes as in the FGK sample if the timescale for abiogenesis is short }(see Section~\ref{sec:results-bioverse}).
A potentially inhabited planet's host star spectral type may thus not be a strong indicator for the credibility of a biosignature detection in the context of the UV~Threshold Hypothesis.

@@

\appendix
\section{\rev{Alternative priors for the abiogenesis rate}}
\label{sec:appendix-flife-prior}
\rev{In the main text, we used a log-uniform prior for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.
Nevertheless, we repeated our semi-analytical analysis for the more optimistic choice of a prior that is \textit{uniform} in $f_\mathrm{life}$.
Figure~\ref{fig:semian_true_evidence_uniform} shows the resulting probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis under this prior.
While the resulting trends are qualitatively similar, the uniform prior yields overall higher probabilities of such a conclusive test for all sample sizes.
}

\rev{
The overall impact differs significantly between M dwarf and FGK dwarf planets.
 While M dwarf planets remain largely unaffected, the fraction of inhabited FGK dwarf planets decreases sharply due to the limited overlap between the time of high UV flux and habitable conditions (compare Figure~\ref{fig:hz_nuv_evo}).
Even under the optimistic assumption that the probability of abiogenesis is unity (\( f_{\text{life}} = 1 \)), the fraction of inhabited planets under the UV Threshold Hypothesis declines rapidly with increasing UV flux thresholds (Figure~\ref{fig:inhabited_dT100}).
}
\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/inhabited_aafo_nuv_thresh_deltatmin100.pdf}
    \caption{\rev{Fraction of inhabited planets as a function of NUV threshold flux for \(\Delta T_{\text{min}} = 100\) Myr. The decrease in inhabited planets is more pronounced for FGK stars, narrowing the viable range of UV Thresholds compared to shorter origins timescales.}}
    \label{fig:inhabited_dT100}
\end{figure}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/surveys_FGKM_dT100.pdf}
    \caption{\rev{Simulated survey results under the assumption of \(\Delta T_{\text{min}} = 100\) Myr. The fraction of inhabited FGK planets is reduced to zero, while the fraction of inhabited M dwarf planets remains largely unaffected.}}
    \label{fig:surveys_dT100}
\end{figure*}

\rev{
The impact of a longer origins timescale on testing the UV Threshold Hypothesis becomes evident when we repeat our example survey ($f_{\text{life}} = 0.8$, $F_{\text{NUV},\text{min}} = \SI{300}{\erg\per\second\per\centi\meter\squared}$) under this assumption (Figure~\ref{fig:surveys_dT100}):
While the survey of M dwarf planets remains largely unaffected, the fraction of inhabited FGK planets is reduced to zero, making a conclusive test of the UV Threshold Hypothesis impossible.
}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/powergrid_dT100.pdf}
    \caption{\rev{Statistical power for different abiogenesis rates and NUV threshold fluxes under the assumption of \(\Delta T_{\text{min}} = 100\) Myr. The longer origins timescale makes a test of the UV Threshold Hypothesis impossible for FGK planets, while M dwarf planets remain viable targets.}}
    \label{fig:powergrid_dT100}
\end{figure*}
\rev{
Figure \ref{fig:powergrid_dT100} demonstrates that this result is independent of the threshold \gls{NUV} flux and the intrinsic abiogenesis rate.
Given the lack of constraints on the timescale of abiogenesis, these findings highlight the advantages of focusing on M dwarf planets when searching for biosignatures.
If life emerges slowly, FGK planets may rarely, if ever, reach the inhabited stage, whereas M dwarf planets remain viable targets for future surveys.
If biosignatures were nevertheless detected on FGK planets despite the constraints imposed by a longer abiogenesis timescale, it would suggest that either the UV Threshold Hypothesis, as formulated here, is incorrect or that the timescale of abiogenesis is shorter than $\mathcal{O}(\SI{100}{\mega\year})$.
}

@@

\end{document}

>>>>

We \rev{first} follow a semi-analytical Bayesian analysis to estimate probabilities of obtaining strong evidence for or against \rev{the UV Threshold Hypothesis}.
Under \rev{this hypothesis} ($H_1$), the probability of an exoplanet having detectable biosignatures is zero if the \rev{\gls{NUV}} irradiation is less than the threshold, and it is equal to the (unknown) probability of \rev{life emerging and persisting}, $f_\mathrm{life}$, if \gls{NUV} exceeds the threshold for a sufficiently long period of time.
Under the null hypothesis ($H_\mathrm{null}$), that probability simply is $f_\mathrm{life}$, that is, it does not correlate with the \gls{UV} flux.

@@

Our simulations find no clear criterion for the credibility of a biosignature detection based on spectral type of the host star, as both FGK and M~dwarf samples show similar maximum past \gls{NUV} flux distributions.
\rev{While there is a significant preference for M~dwarfs in the case of a long timescale for abiogenesis (see Appendix \ref{sec:long-origins}
), the fraction of inhabited planets in the M~dwarf sample drops at similar threshold fluxes as in the FGK sample if the timescale for abiogenesis is short }(see Section~\ref{sec:results-bioverse}).
A potentially inhabited planet's host star spectral type may thus not be a strong indicator for the credibility of a biosignature detection in the context of the UV~Threshold Hypothesis.

@@

\appendix

@@

\section{\rev{Alternative priors for the abiogenesis rate}}
\label{sec:appendix-flife-prior}
\rev{In the main text, we used a log-uniform prior for the probability of life emerging and persisting $f_\mathrm{life}$ in order to reflect our ignorance about this parameter and its order of magnitude.
We consider this a reasonable choice, as it is agnostic about the scale of $f_\mathrm{life}$ and assigns equal prior probability to all orders of magnitude.
Nevertheless, we repeated our semi-analytical analysis for the more optimistic choice of a prior that is \textit{uniform} in $f_\mathrm{life}$.
Figure~\ref{fig:semian_true_evidence_uniform} shows the resulting probabilities of obtaining strong evidence for or against the UV~Threshold Hypothesis under this prior.
While the resulting trends are qualitatively similar, the uniform prior yields overall higher probabilities of such a conclusive test for all sample sizes.
}

\rev{
The overall impact differs significantly between M dwarf and FGK dwarf planets.
 While M dwarf planets remain largely unaffected, the fraction of inhabited FGK dwarf planets decreases sharply due to the limited overlap between the time of high UV flux and habitable conditions (compare Figure~\ref{fig:hz_nuv_evo}).
Even under the optimistic assumption that the probability of abiogenesis is unity (\( f_{\text{life}} = 1 \)), the fraction of inhabited planets under the UV Threshold Hypothesis declines rapidly with increasing UV flux thresholds (Figure~\ref{fig:inhabited_dT100}).
}
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figures/inhabited_aafo_nuv_thresh_deltatmin100.pdf}
    \caption{\rev{Fraction of inhabited planets as a function of NUV threshold flux for \(\Delta T_{\text{min}} = 100\) Myr. The decrease in inhabited planets is more pronounced for FGK stars, narrowing the viable range of UV Thresholds compared to shorter origins timescales.}}
    \label{fig:inhabited_dT100}
\end{figure}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/surveys_FGKM_dT100.pdf}
    \caption{\rev{Simulated survey results under the assumption of \(\Delta T_{\text{min}} = 100\) Myr. The fraction of inhabited FGK planets is reduced to zero, while the fraction of inhabited M dwarf planets remains largely unaffected.}}
    \label{fig:surveys_dT100}
\end{figure*}

\rev{
The impact of a longer origins timescale on testing the UV Threshold Hypothesis becomes evident when we repeat our example survey ($f_{\text{life}} = 0.8$, $F_{\text{NUV},\text{min}} = \SI{300}{\erg\per\second\per\centi\meter\squared}$) under this assumption (Figure~\ref{fig:surveys_dT100}):
While the survey of M dwarf planets remains largely unaffected, the fraction of inhabited FGK planets is reduced to zero, making a conclusive test of the UV Threshold Hypothesis impossible.
}
\begin{figure*}
    \centering
    \includegraphics[width=0.9\linewidth]{figures/powergrid_dT100.pdf}
    \caption{\rev{Statistical power for different abiogenesis rates and NUV threshold fluxes under the assumption of \(\Delta T_{\text{min}} = 100\) Myr. The longer origins timescale makes a test of the UV Threshold Hypothesis impossible for FGK planets, while M dwarf planets remain viable targets.}}
    \label{fig:powergrid_dT100}
\end{figure*}
\rev{
Figure \ref{fig:powergrid_dT100} demonstrates that this result is independent of the threshold \gls{NUV} flux and the intrinsic abiogenesis rate.
Given the lack of constraints on the timescale of abiogenesis, these findings highlight the advantages of focusing on M dwarf planets when searching for biosignatures.
If life emerges slowly, FGK planets may rarely, if ever, reach the inhabited stage, whereas M dwarf planets remain viable targets for future surveys.
If biosignatures were nevertheless detected on FGK planets despite the constraints imposed by a longer abiogenesis timescale, it would suggest that either the UV Threshold Hypothesis, as formulated here, is incorrect or that the timescale of abiogenesis is shorter than $\mathcal{O}(\SI{100}{\mega\year})$.
}
\end{document}
